---
title: "Biostat 203B Homework 2"
subtitle: Due Feb 6 @ 11:59PM
author: Lilly Huang
output: 
  html_document:
    toc: true
    toc_depth: 4 
---

Display machine information for reproducibility:
```{r}
sessionInfo()
```

```{r setup, message=F}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, cache.lazy = FALSE)
library(tidyverse)
library(data.table)
library(lubridate)
```

```{r}
os <- sessionInfo()$running
if (str_detect(os, "Linux")) {
  mimic_path <- "/mnt/mimiciv/1.0"
} else if (str_detect(os, "macOS")) {
  mimic_path <- "/Users/huazhou/Documents/Box Sync/MIMIC/mimic-iv-1.0"
}
```

In this exercise, we use tidyverse (ggpot2, dplyr, etc) to explore the [MIMIC-IV](https://mimic.mit.edu/docs/iv/) data introduced in [homework 1](https://ucla-biostat-203b.github.io/2022winter/hw/hw1/hw1.html) and to build a cohort of ICU stays.

```{r}
# tree -s -L 2 /Users/huazhou/Documents/Box\ Sync/MIMIC/mimic-iv-1.0
system(str_c("tree -s -L 2 ", shQuote(mimic_path)), intern = TRUE)
```

## Q1. `read.csv` (base R) vs `read_csv` (tidyverse) vs `fread` (data.table)

There are quite a few utilities in R for reading plain text data files. Let us test the speed of reading a moderate sized compressed csv file, `admissions.csv.gz`, by three programs: `read.csv` in base R, `read_csv` in tidyverse, and `fread` in the popular data.table package. 

Which function is fastest? Is there difference in the (default) parsed data types? (Hint: R function `system.time` measures run times.)

For later questions, we stick to the tidyverse.
**Solution**

```{r}
admissiondt = "/mnt/mimiciv/1.0/core/admissions.csv.gz"
```
```{r}
system.time(read.csv(admissiondt))
```
```{r}
system.time(read_csv(admissiondt))
```
```{r}
system.time(fread(admissiondt))
```

Fread in the popular data.table package is the fastest.There is a difference in how they read the data. 'Fread' read the data by mapping.However, 'read.csv' read the data line by line.

## Q2. ICU stays

`icustays.csv.gz` (<https://mimic.mit.edu/docs/iv/modules/icu/icustays/>) contains data about Intensive Care Units (ICU) stays. The first 10 lines are
```{r}
system(
  str_c(
    "zcat < ", 
    shQuote(str_c(mimic_path, "/icu/icustays.csv.gz")), 
    " | head"
    ), 
  intern = TRUE
)
```

1. Import `icustatys.csv.gz` as a tibble `icustays_tble`.


```{r}
library(tibble)
```

```{r}
icustays_tble =
  read_csv(str_c(mimic_path, "/icu/icustays.csv.gz"))
```


2. How many unique `subject_id`? Can a `subject_id` have multiple ICU stays? 

**solution:**

```{r}
icustays_tble %>%
  distinct(subject_id) %>%
    nrow() %>%
      str_c(., "unique subject_id")  
```


```{r}
icustays_tble %>%
  distinct(stay_id) %>% 
    nrow() %>%
      str_c(., "unique stay_id")  
```

Yes, since 76540 unique stay_id presents while only 53150 unique subject_id presents,therefor, subject_id can have multipe ICU stays.


3. For each `subject_id`, let's only keep the first ICU stay in the tibble `icustays_tble`.

```{r}
icustays_tble =
  distinct(icustays_tble, subject_id, .keep_all = TRUE)
```



## Q3. `admission` data

Information of the patients admitted into hospital is available in `admissions.csv.gz`. See <https://mimic.mit.edu/docs/iv/modules/core/admissions/> for details of each field in this file. The first 10 lines are
```{r}
system(
  str_c(
    "zcat < ", 
    shQuote(str_c(mimic_path, "/core/admissions.csv.gz")), 
    " | head"
    ), 
  intern = TRUE
)
```

1. Import `admissions.csv.gz` as a tibble `admissions_tble`.

```{r}
admission_tble =
  read_csv(str_c(mimic_path, "/core/admissions.csv.gz"))
```

2. Let's only keep the admissions that have a match in `icustays_tble` according to `subject_id` and `hadmi_id`.

```{r}
admission_tble =
  semi_join(admission_tble, icustays_tble, by = c("subject_id", "hadm_id"))
```


3. Summarize the following variables by graphics. 

- admission year  
```{r}
admyeartb=
ggplot(admission_tble) +
  geom_freqpoly(aes(year(admittime))) +
  labs(x = "admission year", y = "rate",
       title = "Table of Hospital Annully ICU Admission Rate")
admyeartb
```


- admission month  

```{r}
admmontb=
ggplot(admission_tble) +
  geom_density(aes(month(admittime))) +
  labs(x = "admission month", y = "rate",
       title = "Table of Hospital Monthly ICU Admission Rate")
admmontb
```

- admission month day 
```{r}
admmonthdaytb=
ggplot(admission_tble) +
  geom_density(aes(mday(admittime))) +
  labs(x = "admission month day", y = "rate",
       title = "Table of Hospital Month Day ICU Admission Rate")
admmonthdaytb
```

- admission week day  
```{r}
admweekdaytb=
ggplot(admission_tble) +
  geom_bar(stat="count") + aes(weekdays(admittime))
  labs(x = "admission week day", y = "count",
       title = "Table of Hospital Week day ICU Admission Count")
admweekdaytb
```

- admission hour (anything unusual?)  
```{r}
hourtb=
ggplot(admission_tble) +
  geom_bar(stat="count") + aes(hour(admittime))
  labs(x = "admission hour", y = "count",
       title = "Table of Hourly Hospital ICU Admission Count")
hourtb
```


## Q4. `patients` data

Patient information is available in `patients.csv.gz`. See <https://mimic.mit.edu/docs/iv/modules/core/patients/> for details of each field in this file. The first 10 lines are
```{r}
system(
  str_c(
    "zcat < ", 
    shQuote(str_c(mimic_path, "/core/patients.csv.gz")), 
    " | head"
    ), 
  intern = TRUE
)
```

1. Import `patients.csv.gz` (<https://mimic.mit.edu/docs/iv/modules/core/patients/>) as a tibble `patients_tble` and only keep the patients who have a match in `icustays_tble` (according to `subject_id`).
```{r}
patients_tble =
  read_csv(str_c(mimic_path, "/core/patients.csv.gz"))
```


```{r}
patients_tble =
  semi_join(patients_tble, icustays_tble, 
      by = c("subject_id"))
```


2. Summarize variables `gender` and `anchor_age`, and explain any patterns you see.

```{r}
genderplot=
ggplot(patients_tble) +
  geom_boxplot(aes(x=gender,y=anchor_age)) +
  labs(x = "gender", y = "anchor_age",
       title = "The Anchor Age Of Patients In The ICU According To The Gender")
genderplot

```

The pattern is that male tend to have a slightly lower average anchor age than female patients in the ICU. The ranges of the anchor age for both gender are very similar.



## Q5. Lab results

`labevents.csv.gz` (<https://mimic.mit.edu/docs/iv/modules/hosp/labevents/>) contains all laboratory measurements for patients. The first 10 lines are
```{r}
system(
  str_c(
    "zcat < ", 
    shQuote(str_c(mimic_path, "/hosp/labevents.csv.gz")), 
    " | head"
    ), 
  intern = TRUE
)
```
`d_labitems.csv.gz` is the dictionary of lab measurements. 
```{r}
system(
  str_c(
    "zcat < ", 
    shQuote(str_c(mimic_path, "/hosp/d_labitems.csv.gz")), 
    " | head"
    ), 
  intern = TRUE
)
```

1. Find how many rows are in `labevents.csv.gz`.

```{r}
labevents=
read_csv ("/mnt/mimiciv/1.0/hosp/labevents_filtered_itemid.csv.gz")
  count(labevents)
```


2. We are interested in the lab measurements of creatinine (50912), potassium (50971), sodium (50983), chloride (50902), bicarbonate (50882), hematocrit (51221), white blood cell count (51301), glucose (50931), magnesium (50960), and calcium (50893). Retrieve a subset of `labevents.csv.gz` only containing these items for the patients in `icustays_tble` as a tibble `labevents_tble`. 

    Hint: `labevents.csv.gz` is a data file too big to be read in by the `read_csv` function in its default setting. Utilize the `col_select` and `lazy` options in the `read_csv` function to reduce the memory burden.
    
```{r}
labevents_tble = 
  read_csv("/mnt/mimiciv/1.0/hosp/labevents_filtered_itemid.csv.gz",
          col_select = c("subject_id", "itemid"), lazy = TRUE) %>%
  filter(itemid == "50912" & itemid == "50971" 
         & itemid == "50983" & itemid == "50902" & itemid == "50882"
         & itemid == "51221" & itemid == "51301" & itemid == "50931"
         & itemid == "50960" & itemid == "50893") %>%
  semi_join(icustays_tble, by = "subject_id")
```


3. Further restrict `labevents_tble` to the first lab measurement during the ICU stay.

```{r}
labevents_tble=
  distinct(labevents_tble, subject_id, .keep_all = TRUE)
```

4. Summarize the lab measurements by appropriate numerics and graphics. 

```{r}

```

## Q6. Vitals from charted events

`chartevents.csv.gz` (<https://mimic.mit.edu/docs/iv/modules/icu/chartevents/>) contains all the charted data available for a patient. During their ICU stay, the primary repository of a patient’s information is their electronic chart. The `itemid` variable indicates a single measurement type in the database. The `value` variable is the value measured for `itemid`. The first 10 lines of `chartevents.csv.gz` are
```{r}
system(
  str_c(
    "zcat < ", 
    shQuote(str_c(mimic_path, "/icu/chartevents.csv.gz")), 
    " | head"), 
  intern = TRUE
)
```
`d_items.csv.gz` (<https://mimic.mit.edu/docs/iv/modules/icu/d_items/>) is the dictionary for the `itemid` in `chartevents.csv.gz`. 
```{r}
system(
  str_c(
    "zcat < ", 
    shQuote(str_c(mimic_path, "/icu/d_items.csv.gz")), 
    " | head"), 
  intern = TRUE
)
```

1. We are interested in the vitals for ICU patients: heart rate (220045), mean non-invasive blood pressure (220181), systolic non-invasive blood pressure (220179), body temperature in Fahrenheit (223761), and respiratory rate (220210). Retrieve a subset of `chartevents.csv.gz` only containing these items for the patients in `icustays_tble` as a tibble `chartevents_tble`.

```{r}
chartevents_tble =
  read_csv("/mnt/mimiciv/1.0/icu/chartevents_filtered_itemid.csv.gz",
          col_select = c("subject_id", "itemid"), lazy = TRUE) %>%
  filter(itemid == "220045" & itemid == "220181" 
         & itemid == "220179" & itemid == "223761" & itemid == "220210")
  semi_join(icustays_tble, by = "subject_id")
```


2. Further restrict `chartevents_tble` to the first vital measurement during the ICU stay. 

```{r}
chartevents_tble=
  distinct(chartevents_tble, subject_id, .keep_all = TRUE)
```


3. Summarize these vital measurements by appropriate numerics and graphics.

```{r}

```


## Q7. Putting things together

Let us create a tibble `mimic_icu_cohort` for all ICU stays, where rows are  

- first ICU stay of each unique adult (age at admission > 18)

and columns contain at least following variables  

- all variables in `icustays.csv.gz`  
- all variables in `admission.csv.gz`  
- all variables in `patients.csv.gz`  
- first lab measurements during ICU stay  
- first vital measurements during ICU stay
- an indicator variable `thirty_day_mort` whether the patient died within 30 days of hospital admission (30 day mortality)

```{r}
mimic_icu_cohort=
  icustays_tble %>%
  left_join(admission_tble, by = c("subject_id","hadm_id")) %>%
  left_join(patients_tble, by = c("subject_id")) %>%
  left_join(labevents_tble, by = c("subject_id")) %>%
  left_join(chartevents_tble, by = c("stay_id")) %>%
  mutate(age_hadm = anchor_age + year(admittime) - anchor_year) %>%
  filter(age_hadm > 18) %>%
  mutate(hadm_death = ifelse(is.na(deathtime), Inf, deathtime -admittime)) %>%
  mutate((thirty_day_mort) = hadm_death) %>%
  print(width = Inf)
  write_rds("mimic_icu_cohort.rds")
```



## Q8. Exploratory data analysis (EDA)

Summarize following information using appropriate numerics or graphs.

- `thirty_day_mort` vs demographic variables (ethnicity, language, insurance, marital_status, gender, age at hospital admission)

- `thirty_day_mort` vs first lab measurements

- `thirty_day_mort` vs first vital measurements

- `thirty_day_mort` vs first ICU unit

